Regression Cube Analysis of Cohort Study Data

### 2015-06-20 draft.tex
covered security issues

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
%%
\begin{enumerate}
	%%
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	%%
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
	%%
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
\magenta{
The \emph{Regression Cube} is designed to provide overview visualizations to support hypothesis generation.
%%
Hence, it is associated with step 1 and 2 of the epidemiological workflow presented in Section~\ref{sec:EpidemiologicalWorkflow}.
%%
Relationships observed using such techniques are subject of detailed statistical testing by statisticians with background in epidemiology using statistical processors, such as SPSS or STATA.
}
% \design{
% This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
% }
\subsection{Iterative Design Based on Expert Feedback}
%%
\design{
%%
The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
%%
The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
%%
It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
%%
That is the focus of this work.
%%
% These target features often indicate the presence of the investigated disease.
%%
The domain experts wanted to \emph{model knowledge} about the investigated condition, such as confounding features (e.g., age or gender).
%%
For explorative analysis, they preferred an approach, which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
%%
Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
%%
This aspect is considered later in the Section~\ref{sec:SystemDesign}.
%%
%In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
}
%%
%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
%Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
%%
%Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
%In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
%These target features often indicate the presence of the investigated disease.
%%

@@

>>>>

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
We concluded that the model quality is insufficient to characterize back pain.
The difference of the presented approach towards our previous and related work is twofold:
\begin{enumerate}
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
\section{Regression Cube Analysis of Cohort Study Data}
\magenta{
The \emph{Regression Cube} is designed to provide overview visualizations to support hypothesis generation.
Hence, it is associated with step 1 and 2 of the epidemiological workflow presented in Section~\ref{sec:EpidemiologicalWorkflow}.
Relationships observed using such techniques are subject of detailed statistical testing by statisticians with background in epidemiology using statistical processors, such as SPSS or STATA.
% \design{
% This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
% }
\subsection{Iterative Design Based on Expert Feedback}
The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
That is the focus of this work.
% These target features often indicate the presence of the investigated disease.
The domain experts wanted to \emph{model knowledge} about the investigated condition, such as confounding features (e.g., age or gender).
For explorative analysis, they preferred an approach, which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
%Concerns regarding security issues the web-based approach were raised by domain experts due to the 
Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
Technical measures taken to enable a secure transfer and storage are described later in the Section~\ref{sec:SystemDesign}.
%In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
%Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%These target features often indicate the presence of the investigated disease.

@@

\magenta{
This architecture is also prone to security issues, such as the storage of confidential data, especially in the epidemiological context.
Therefore, we have to incorporate technical measures to ensure a secure workflow.
